Polynomial phase estimation by phase unwrapping

### 2013-04-15 paper.tex
minor paper mod

%\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
\documentclass[journal]{IEEEtran}

@@

An analysis of the HAF estimator under the assumption that the noise process $\{X_n\}$ is white and Gaussian and without the high SNR assumption has been conducted in~\cite{Porat_asympt_HAF_DPT_1996} (see Lemma~1 and the remark that follows in~\cite{Porat_asympt_HAF_DPT_1996}).  To date, the literature does not contain a corresponding analysis of any estimator based on phase unwrapping.  This has apparently led to the misconception that phase unwrapping estimators are only accurate when the noise is `small' in some sense.  This misconception may have been motivated by early papers such as~\cite{Tretter1985,Kay1989} where it was observed that the noise occurring in the phase of the signal only looks `Gaussian' when the complex noise producing it has small variance.  The results in this paper make it clear that phase unwrapping estimators can be accurate at arbitrarily low SNR.  The information contained in the amplitude is not essential for accurate estimation of polynomial phase signals, and can be discarded with only small loss in accuracy.  For example, in the case that $\{X_n\}$ is white and Gaussian, our results show that, asymptotically in $N$, the ratio of the variance of the LSU estimator to the Cramer-Rao lower bound (CRB) converges to $1$ for high SNR, and to $\tfrac{2}{3}\pi$ for low SNR.  Related results have been presented in~\cite{Quinn2009_dasp_phase_only_information_loss}.  This factor of $\tfrac{2}{3}\pi$ for low SNR is better than popular estimators such as the HAF and CPF, for which the ratio of the variance and the CRB diverges as the SNR become small~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.

@@

Figure~\ref{plot:m3small} shows the sample MSE of the highest order coefficient $\mu_3$ for a polynomial phase signal of order $m=3$.  All the estimators perform poorly at low SNR until a `threshold' is reached, after which the variance is close to the Cramer-Rao bound (CRB)~\cite{Peleg1991_CRB_PPS_1991} indicated by the solid line.  The asymptotic variance of the LSU estimator predicted by Theorem~\ref{thm:asymp_proof} is displayed by the dashed line.  Provided the SNR is large enough (so that the `threshold' is avoided) the sample MSE of the LSU estimator is close to that predicted by Theorem~\ref{thm:asymp_proof}.  Using Theorem~\ref{thm:asymp_proof} we can predict that, asymptotically in $N$, the ratio between the CRB and the variance of the LSU estimator will approach $1$ for high SNR and $\tfrac{2}{3}\pi$ for low SNR~(see Lemma~3 and the subsequent discussion in~\cite{Quinn2009_dasp_phase_only_information_loss}).  Contrastingly, the ratio between the CRB and variance of the HAF, PHAF and CPF estimators is known to diverge as the SNR becomes small~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.  %For this reason we can expect the LSU estimator to be more accurate at low SNR for sufficiently large $N$.

>>>>

\documentclass[draftcls, onecolumn, 11pt]{IEEEtran}
%\documentclass[journal]{IEEEtran}

@@

An analysis of the HAF estimator under the assumption that the noise process $\{X_n\}$ is white and Gaussian and without the high SNR assumption has been conducted in~\cite{Porat_asympt_HAF_DPT_1996} (see Lemma~1 and the remark that follows in~\cite{Porat_asympt_HAF_DPT_1996}).  To date, the literature does not contain a corresponding analysis of any estimator based on phase unwrapping.  This has apparently led to the misconception that phase unwrapping estimators are only accurate when the noise is `small' in some sense.  This misconception may have been motivated by early papers such as~\cite{Tretter1985,Kay1989} where it was observed that the noise occurring in the phase of the signal only looks `Gaussian' when the complex noise producing it has small variance.  The results in this paper make it clear that phase unwrapping estimators can be accurate at arbitrarily low SNR.  The information contained in the amplitude is not essential for accurate estimation of polynomial phase signals, and can be discarded with only small loss in accuracy.  For example, in the case that $\{X_n\}$ is white and Gaussian, our results show that, asymptotically in $N$, the ratio of the variance of the LSU estimator to the Cramer-Rao lower bound (CRB) converges to $1$ for high SNR, and to $\tfrac{2}{3}\pi$ for low SNR.  Related results have been presented in~\cite{Quinn2009_dasp_phase_only_information_loss}.  This factor of $\tfrac{2}{3}\pi$ for low SNR is better than popular estimators such as the HAF and CPF, for which the ratio of the variance to the CRB diverges as the SNR becomes small~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.

@@

Figure~\ref{plot:m3small} shows the sample MSE of the highest order coefficient $\mu_3$ for a polynomial phase signal of order $m=3$.  All the estimators perform poorly at low SNR until a `threshold' is reached, after which the variance is close to the Cramer-Rao bound (CRB)~\cite{Peleg1991_CRB_PPS_1991} indicated by the solid line.  The asymptotic variance of the LSU estimator predicted by Theorem~\ref{thm:asymp_proof} is displayed by the dashed line.  Provided the SNR is large enough (so that the `threshold' is avoided) the sample MSE of the LSU estimator is close to that predicted by Theorem~\ref{thm:asymp_proof}.  Using Theorem~\ref{thm:asymp_proof} we can predict that, asymptotically in $N$, the ratio of the variance of the LSU estimator to the CRB will approach $1$ for high SNR and $\tfrac{2}{3}\pi$ for low SNR~(see Lemma~3 and the subsequent discussion in~\cite{Quinn2009_dasp_phase_only_information_loss}).  Contrastingly, the ratio of the variance of the HAF, PHAF and CPF estimators to the CRB is known to diverge as the SNR becomes small~\cite{Porat_asympt_HAF_DPT_1996,Oshea_cpf_2004,Djurovic_haf_cpf_2012}.  %For this reason we can expect the LSU estimator to be more accurate at low SNR for sufficiently large $N$.
